Self Regulation Ontology Retest Analyses

### 2018-04-02 code/SRO_Retest_Analyses.Rmd
added data quality checks

## Demographics reliability

@@

## Effect of delays between the two measurements

@@

Make data frame with difference between two scores for each measure for each subject. Since the scores for different measures are on different scales for comparability the difference scores are normalized (ie demeaned and dividied by the sd of the difference score distribution.) Note that the variance of the difference score distribution accounts for the variance in both time points by summing them. Normalization equates the means of each difference score distribution to 0 which would mask any meaningful change between the two time points but the analysis here does not interpret the mean of the difference score distributions but is interested in its relation to the days between completion. We check if the variables show systematic differences between the two points later and can investigate whether

Here we check if the difference is larger the longer the delay
What is the threshold for a meaningful difference between the two points for each variable? 
For standardized difference scores does it cross 2 on the y axis?

@@

```{r}
t1_2_difference %>%
  mutate(sig_days_effect = ifelse(dv %in% sig_days_effect$dv, 1, 0))%>%
  arrange(-sig_days_effect) %>%
  ggplot()+
  stat_smooth (aes(as.numeric(days_btw), abs(difference), 
                  group=factor(dv, levels=unique(dv[order(sig_days_effect)]), ordered=TRUE), 
                  color=factor(sig_days_effect, levels = c(1,0))),geom="line", alpha=0.5, method='lm')+
  geom_smooth(aes(as.numeric(days_btw), abs(difference)), method='lm', color = "black", se=FALSE)+
  theme(legend.title = element_blank())+
  xlab('Days between completion')+
  ylab('Scaled difference score')+
  scale_color_discrete(breaks=c(0,1),
                       labels=c("NS", "Sig"))
```

@@

@@

`Plot difference score distributions for the measures with significant differences as examples`

```{r}
t1_2_difference %>% 
  filter(dv == sig_days_effect$dv[1]) %>%
  mutate(days_btw = as.numeric(days_btw),
         days_bins = ifelse(days_btw < quantile(days_btw, probs=0.33), "short", 
                                                ifelse(days_btw>quantile(days_btw, probs=0.66), "large", "medium"))) %>%
  ggplot(aes(factor(days_bins, levels = c("short", "medium", "large")), difference))+
  geom_boxplot()+
  xlab('delay between two measurements')
```

```{r}
t1_2_difference %>% 
  filter(dv == sig_days_effect$dv[4]) %>%
  mutate(days_btw = as.numeric(days_btw),
         days_bins = ifelse(days_btw < quantile(days_btw, probs=0.33), "short", 
                                                ifelse(days_btw>quantile(days_btw, probs=0.66), "large", "medium"))) %>%
  ggplot(aes(factor(days_bins, levels = c("short", "medium", "large")), difference))+
  geom_boxplot()+
  xlab('delay between two measurements')
```

@@

@@

  mod_lit = lm(retest_reliability ~ lit_predictor, data=sampled_df)
  mod_boot = lm(retest_reliability ~ mean_icc, data=sampled_df)

@@

>>>>

## Data quality checks

### Demographics reliability

@@

### Effect of delays between the two measurements

@@

Make data frame with difference between two scores for each measure for each subject. Since the scores for different measures are on different scales for comparability the difference scores are normalized (ie demeaned and dividied by the sd of the difference score distribution.) Note that the variance of the difference score distribution accounts for the variance in both time points by summing them. Normalization equates the means of each difference score distribution to 0 which would mask any meaningful change between the two time points but the analysis here does not interpret the mean of the difference score distributions but is interested in its relation to the days between completion. We check if the variables show systematic differences between the two points later.

Here we check if the difference is larger the longer the delay

@@

```{r}
t1_2_difference %>%
  mutate(sig_days_effect = ifelse(dv %in% sig_days_effect$dv, 1, 0))%>%
  arrange(-sig_days_effect) %>%
  ggplot()+
  stat_smooth (aes(as.numeric(days_btw), abs(difference), 
                  group=factor(dv, levels=unique(dv[order(sig_days_effect)]), ordered=TRUE), 
                  color=factor(sig_days_effect, levels = c(1,0))),geom="line", alpha=0.5, method='lm')+
  geom_smooth(aes(as.numeric(days_btw), abs(difference)), method='lm', color = "black", se=FALSE)+
  theme(legend.title = element_blank())+
  xlab('Days between completion')+
  ylab('Scaled difference score')+
  scale_color_discrete(breaks=c(0,1),
                       labels=c("NS", "Sig"))

ggsave('DaysBtwEffect.jpg', device = "jpeg", path = "/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/output/figures/", width = 6, height = 4, units = "in", limitsize = FALSE, dpi = 100)
```

@@

Do the variables that show significant dependence on the delay between the completion times have systematically lower/higher reliability? Looking at the reliability point estimate over the size of the delay effect. There are very few variables to compare anyway but regardless doesn't look conclusive/concerning.

@@

### Overlapping survery questions

Some surveys have overlapping questions. Do these correlate within and across sessions?

First determine the overlapping questions.

```{r}
tmp = read.csv(gzfile(paste0(retest_data_path, 'items.csv.gz')))
tmp = tmp %>% 
  filter(worker == 's005') %>% 
  select(item_ID, item_text) %>% 
  mutate(item_text = trimws(as.character(item_text))) %>%
  unite(item, c("item_ID", "item_text"), sep = "___")

comb = as.data.frame(t(combn(unique(tmp$item),2)))

duplicate_items = comb %>% 
  filter(grepl('dospert', V1)==FALSE) %>%
  filter(grepl('selection_optimization', V1)==FALSE) %>%
  filter(grepl('sensation_seeking', V1)==FALSE) %>%
  separate(V1, c("item1_ID", "item1_text"), sep="___") %>%
  separate(V2, c("item2_ID", "item2_text"), sep="___") %>%
  mutate(similarity = levenshteinSim(item1_text, item2_text)) %>%
  filter(similarity>0.8) %>%
  select(item1_ID, item2_ID, item1_text, item2_text)

duplicate_items
```

```{r warning=FALSE, message=FALSE}
#surveys to read in
extract_items = c('worker',unique(with(duplicate_items, c(item1_ID, item2_ID))))

#correlations to compute:
#item1_t1 - item2_t1, 
#item1_t2 - item2_t2, 
#item1_t1 - item2_t2, 
#item1_t2 - item2_t1

duplicate_items_data_t1 = read.csv(paste0(test_data_path, 'subject_x_items.csv'))
duplicate_items_data_t2 = read.csv(paste0(retest_data_path, 'subject_x_items.csv'))

duplicate_items_data_t1 = duplicate_items_data_t1 %>%
  filter(worker %in% duplicate_items_data_t2$worker) %>%
  select(extract_items)

duplicate_items_data_t2=duplicate_items_data_t2 %>%
  filter(worker %in% duplicate_items_data_t1$worker) %>%
  select(extract_items)

duplicate_items = duplicate_items %>%
  mutate(t1_t1_cor = NA,
         t2_t2_cor = NA,
         t1_t2_cor = NA,
         t2_t1_cor = NA,
         t1_t1_polycor = NA,
         t2_t2_polycor = NA,
         t1_t2_polycor = NA,
         t2_t1_polycor = NA)

for(i in 1:nrow(duplicate_items)){
  duplicate_items$t1_t1_cor[i] = abs(cor(duplicate_items_data_t1[,c(duplicate_items$item1_ID[i])],
                                  duplicate_items_data_t1[,c(duplicate_items$item2_ID[i])]))

  duplicate_items$t2_t2_cor[i] = abs(cor(duplicate_items_data_t2[,c(duplicate_items$item1_ID[i])],
                                  duplicate_items_data_t2[,c(duplicate_items$item2_ID[i])]))

  duplicate_items$t1_t2_cor[i] = abs(cor(duplicate_items_data_t1[,c(duplicate_items$item1_ID[i])],
                                  duplicate_items_data_t2[,c(duplicate_items$item2_ID[i])]))

  duplicate_items$t2_t1_cor[i] = abs(cor(duplicate_items_data_t2[,c(duplicate_items$item1_ID[i])],
                                  duplicate_items_data_t1[,c(duplicate_items$item2_ID[i])]))

  duplicate_items$t1_t1_polycor[i] = abs(polychoric(data.frame(duplicate_items_data_t1[,c(duplicate_items$item1_ID[i])],
                      duplicate_items_data_t1[,c(duplicate_items$item2_ID[i])]))$rho[2])

  duplicate_items$t2_t2_polycor[i] = abs(polychoric(data.frame(duplicate_items_data_t2[,c(duplicate_items$item1_ID[i])],
                      duplicate_items_data_t2[,c(duplicate_items$item2_ID[i])]))$rho[2])

  duplicate_items$t1_t2_polycor[i] = abs(polychoric(data.frame(duplicate_items_data_t1[,c(duplicate_items$item1_ID[i])],
                      duplicate_items_data_t2[,c(duplicate_items$item2_ID[i])]))$rho[2])

  duplicate_items$t2_t1_polycor[i] = abs(polychoric(data.frame(duplicate_items_data_t2[,c(duplicate_items$item1_ID[i])],
                      duplicate_items_data_t1[,c(duplicate_items$item2_ID[i])]))$rho[2])
}

duplicate_items
```

```{r}
summary(duplicate_items$t1_t1_polycor)
summary(duplicate_items$t2_t2_polycor)
summary(duplicate_items$t1_t2_polycor)
summary(duplicate_items$t2_t1_polycor)
```

```{r echo=FALSE}
rm(tmp, comb)
```

@@

```{r}
summary(lm(retest_reliability ~ task,rel_comp))
```

@@

  mod_lit = lm(retest_reliability ~ lit_predictor+scale(sample_size), data=sampled_df)
  mod_boot = lm(retest_reliability ~ mean_icc+scale(sample_size), data=sampled_df)

@@

```{r}
with(comp_lit_pred_out, t.test(r2_lit, r2_boot))
```
